% Copyright 2017 Beckman Coulter, Inc.
%
% Permission is hereby granted, free of charge, to any person
% obtaining a copy of this software and associated documentation files
% (the "Software"), to deal in the Software without restriction,
% including without limitation the rights to use, copy, modify, merge,
% publish, distribute, sublicense, and/or sell copies of the Software,
% and to permit persons to whom the Software is furnished to do so,
% subject to the following conditions:
%
% The above copyright notice and this permission notice shall be
% included in all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
% EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
% MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND
% NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS
% BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN
% ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN
% CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
% SOFTWARE.

\documentclass[letterpaper,11pt,oneside,final]{article}
\usepackage{sagian}

\begin{document}

\title {\cpp Coding Standard}
\author {Chris Payne \and Bob Burger}
\date {\copyright\ 2017 Beckman Coulter, Inc.
  Licensed under the \href{https://opensource.org/licenses/MIT}{MIT License}.}
\maketitle

\section* {Purpose}

This document describes the coding standard for \cpp code.

\section* {Counting Standard}

Count each physical line as one \emph{line of code} (LOC). Blank lines
and comment-only lines will be included. A blank line should appear
between each top-level definition. Be consistent about what is on each
physical line.

\section* {Identifiers}

Use meaningful names---this can be very difficult. Identifiers should
correspond directly to the names used in the design
documents. Identifiers should use upper camel casing.

\section* {Comments}

Comments should be clear, concise, and written in U.S. English.

Avoid using comments that say what the code says.

\section* {Formatting}

\cpp code should be indented following the pattern established by
Visual Studio. Indenting should be set to smart. Tab size and indent
size should be set to 2. Insert spaces rather than using tabs.

Line length should be limited to 79 characters. Lines longer than 79
characters may be tolerated.

Lines should end in LF, not CRLF.

Tab characters should not appear in \cpp code. Their presence in a
file makes the indentation depend on the software used to view the
file.

\end{document}
